Trajectory model of passive kelp dispersal

### 2020-06-29 thesis_chapter_3.Rmd
Update thesis_chapter_3.Rmd

Other aspects such as buoyancy and drag also play a role in determining the trajectory and rate of transport for surface floating material. However, the past research on macroalgal trajectory has not investigated these factors which have been shown to be important aspects of trajectory for other materials such as icebergs, marine craft and microplastics. The "sail" area of an object floating at sea is the surface area exposed to the wind which results in air form drag, while the area of the object below the surface of the water is exposed to surface currents which result in water form drag. Furthermore, past research conducted within the maritime industry has shown that the size and shape of a vessel determine the relative importance of waves or wind as influencers of trajectory, as well as orientation of the object. If the length of the object is longer than the significant wave height then waves will be the primary influencer of trajectory while the opposite is true for smaller objects where the effect of waves is regarded as negligible [@breivik2011; @griffin2017]. There are considerable gaps in the literature with regards to macroalgae trajectory such as buoyancy, air form drag, water form drag, epibiont load, growth, fragmentation, size, shape and orientation which have not been investigated numerically in the past.

In order to determine the coefficients needed for the transport model, a series of experiments are proposed. These experiments will aim to determine aspects of buoyancy, epiphyte growth rates and transport rate, all of which have been regarded as important coefficients needed for an accurate model. The experiments will be perfomred numerically to determine the main influencers of passive solitary kelp dispersal.

>>>>

Other aspects such as buoyancy and drag also play a role in determining the trajectory and rate of transport for surface floating material. However, the past research on macroalgal trajectory has not investigated these factors which have been shown to be important aspects of trajectory for other materials such as icebergs, marine craft and microplastics. The "sail" area of an object floating at sea is the surface area exposed to the wind which results in air form drag, while the area of the object below the surface of the water is exposed to surface currents which result in water form drag. Furthermore, past research conducted within the maritime industry has shown that the size and shape of a vessel determine the relative importance of waves or wind as influencers of trajectory, as well as orientation of the object. If the length of the object is longer than the significant wave height then waves will be the primary influencer of trajectory while the opposite is true for smaller objects where the effect of waves is regarded as negligible [@breivik2011; @griffin2017].

Most of the past studies have focused on aspects of raft longevity, with various raft times for a handful of species estimated. Although this is an area that requires more research, other factors that are important in the dispersal of macroalgae have been given comparatively little attention in recent years. An example is the relative importance of surface currents, winds, and waves in the dispersal of marine algae. In addition water- and -air form drag has not been considered before in relation to passive macroalgal dispersal. Drag coefficients related to wind and surface current have been shown to be important properties to consider when estimating trajectory and forecasting drift for search and sea rescue operations. Drag is ultimately determined by the size and shape of the object which are properties that vary considerably with macroalgae not only across species but also across environments.

Langrangian trajectory modeling is a useful tool, however, the various models available often assume the object to be a spherical particle, which is obviously not the case for both biological and anthropogenic forms of marine debris in nature. To accurately determine the trajectory of marine macroalgae these aspects need to be taken into account. Past research by @plourde1999 and @breivik2008 have provided estimates of drag for various objects based on experimental work, which consisted of both direct and indirect methods of the Leeway model. However, this work does not consider biological material such as macroalgae. Although drag estimates do exist, these have been calculated for macroalgae not detached from the substratum and are regarded as fixed-point estimates.

Methods used in the field of chemical engineering provide an alternative approach to estimating drag of macroalgae by means of formulae to calculate drag based on shape.

There are considerable gaps in the literature with regards to macroalgae trajectory such as buoyancy, air form drag, water form drag, epibiont load, growth, fragmentation, size, shape and orientation which have not been investigated numerically in the past.  In order to determine the coefficients needed for the transport model, a series of experiments are proposed. These experiments will aim to determine aspects of buoyancy, epiphyte growth rates and transport rate, all of which have been regarded as important coefficients needed for an accurate model. The experiments will be perfomred numerically to determine the main influencers of passive solitary kelp dispersal.
